Sexual selection improves population fitness: a meta-analysis

### 2018-11-26 Response_to_Nature_Comms.Rmd

>*Another major problem I have with the manuscript concerns interpretation of the very intriguing finding that the response to manipulation of sexual selection was stronger for female traits compared to male traits. I’m confused by the authors’ explanation: do they assume sexual selection acted directly on females, and not only indirectly, via males? Only then things like mother to daughter heritability, or hard selection on females, should matter.*

Leave this for Luke

Good sugggestion!

>*Perhaps the effect on females is indeed direct, and results from stress imposed by polygamous treatment, which magnifies direct, hard selection on females? This would be an important finding, and perhaps the authors could test it with their dataset by contrasting middle-class-neighborhood-like studies from those which allowed for female evolution. But if correct, this explanation is not exactly the effect of sexual selection, but rather enhanced selection of females due to enhanced (male induced) stress, so the interpretation of results should change.*

Answer

**Other comments:**  
>*l. 30 – reviews on sexual conflict are OK to cite here, but there are empirical papers actually demonstrating correlation between male sexual selected traits (Harano et al. 2011; Plesnar et al. 2014) which should also be cited.*

Thank you, we now cite those papers as suggested in line 31. Additionally, they are especially pertenant to experimental evolution studies using species such as the bulb mite (*Rhizoglyphus robini*)[@Plesnar_2014] and flour beetles (*Gnatocerus cornutus*)[@Harano_2010].

>*l. 151 – the authors discuss beneficial effects of sexual selection on direct fitness measures such as reproductive success or offspring viability, but estimates for both of these measures actually overlapped zero! Perhaps joint analysis of direct fitness measures, as I suggested above, could support this conclusion, but currently this is an overstatement.*

The reviewer is correct in that the 95 % CI overlap zero for the three direct fitness components (extinction rate, offspring viability and reproductive success) have 95 % CI overlapping zero. Additionally, our analysis reveals that the predicted effect size for direct fitness components has 95 % CIs that overlap zero for Bayesian, but not REML models (REML $\beta$ = 0.15, 95% CIs: 0.033 to 0.27; Bayesian $\beta$ = 0.15, 95% CIs: -0.058 to 0.34, $BF_{>0}$ = 4.9). In this respect we have tapered and clarified our statement that we find some evidence that sexual selection increases mean values of traits directly related to fitness. However, we retain our statement that sexual selection generally elevates the mean of fitness components directly related to fitness as a two-way 95 % confidence limit is an arbitrary cut-off and the Bayes Factors suggest that their is a high likelihood that the effect size for direct fitness components is greater than zero.

**In the discussion the authors say they included the number of experimental evolution generations, but I could not find this information in methods.**

It is correct that we did not use the number of generations in our main analysis. Line 281 of the manuscript (Methods section) states that the number of generations. In the supplementary material we investigate the role of number of generations (the lineage is evolved for) has on fitness in the section *Meta-analyses including many moderator variables* as well as in Figure S10 and S11. Generations (specifically, log-Generations) had minimal effect on the effect sizes and was removed for the models that appeared in the main text.

**Fig. S1 is not referred to in the main texts, is it different from Fig. 2, except that the latter contains predicted average values for fitness components?**

No difference... kind of there for documentation on how we created it (transparency). What do you think we should do about this Luke?

________________________________________________________

## Comments from Reviewer 2

**This meta-analysis investigated the consequences of sexual selection experiments on trait mean and variance (comparing sexually selected groups vs control groups). Overall, the authors observe sexual selection usually increase mean and reduce variance (especially in females). The authors conclude sexual selection's benefits outweigh its detrimental effects. This meta-analysis is extremely well conducted (the use of both likelihood-based and Bayesian models for robustness), and although I am not an expert on this topic, I really enjoyed reading it, and it was very clear. Especially, I am impressed with the detailed supplement which showed the code and analysis. However, I have several comments regarding their analysis, which will increase the robustness of results and thus conclusions.**

Many thanks for your time and attention, and for the valuable feedback.

**1 --- The use of Hedges' $g$. I understand that Hedges' $g$ was probably used because it can take interval measurements as well as ratio measurements (lnCVR can only take ratio measurements). However, $g$ cannot really deal well with heterogeneity between two groups (i.e. experimental and control groups having different variances). This is why Hedges' $g$ (Cohen's $d$) was criticized earlier.**

Answer

**Osenberg, C. W., O. Sarnelle, and S. D. Cooper. 1997. Effect size in ecological experiments: the application of biological models in meta-analysis. \textit{American Naturalist} 150:798-812.**

Answer

**As a response, they come up with log response ratio (lnRR) - see**

**Hedges, L. V., J. Gurevitch, and P. S. Curtis. 1999. The meta-analysis of response ratios in experimental ecology. \textit{Ecology} 80:1150-1156.**

**I recommend that the authors use lnRR for their effect size for mean comparison as well as Hedge's $g$ to see the robustness of their conclusions for the mean.**

Reviewer 2 has rightly pointed out that the use of Hedge's *g* has been critisized for its ability to account for heterogeneity between the control and treatment groups **CITE**. We extended our meta-analysis by including the log response ratio (lnRR) as well as a modified version of Hedge's g that attempts to account for this heterogeneity, this modified statistic (**CITE**) can be derived from the ``metafor`` package for meta-analysis, where it is named SMDH (standardized mean difference with heteroscedastic population variances in the two groups). There are several key findings from the use of lnRR and SMDH:

*Firstly*, Hedge's g and SMDH are highly correlated. This suggests that unequal variance between the treatment and control groups (heteroscedasticity) is unlikely to impact the effect size of Hedge's g

```{r, warning=F}
#This function writes in the correlation value
panel.cor <- function(x, y, digits=2, prefix="", cex.cor, ...)
{
  usr <- par("usr"); on.exit(par(usr))
  par(usr = c(0, 1, 0, 1))
  r <- abs(cor(x, y, use="pairwise.complete.obs"))
  txt <- format(c(r, 0.123456789), digits=digits)[1]
  txt <- paste0(prefix, txt)
  if(missing(cex.cor)) cex.cor <- 0.8/strwidth(txt)
  text(0.5, 0.5, txt, cex = cex.cor * r)
}

es.correlation.df <- lnRR.data %>% left_join(SMD.data %>% rename(`Hedge's g` = SMD)) %>% left_join(SMDH.data)

pairs(es.correlation.df %>% select(`Hedge's g`, lnRR, SMDH), lower.panel=panel.smooth, upper.panel=panel.cor)
```

*Secondly*, The calculation of lnRR and SMDH can only be calculated for effect sizes where the study reported means, standard deviations and sample sizes (352 out of 459). This is a limitation as in our analysis we investigate the effects of sex and stress on a further restricted dataset (direct and indirect fitness traits only). Specifically, fewer effect sizes from stressful conditions are able to be calculated; making Hedge's g an advantageous effect size in comparison to lnRR.

```{r}
predictions.complete2 %>% select(Sex, Environment, `Hedge's g sample size` = n) %>% 
  left_join(predictions.lnRR %>% select(Sex, Environment, `lnRR sample size` = n), by = c("Sex", "Environment")) %>% pander()
```

*Thirdly*, if we conduct a meta-analysis using lnRR as the effect size we obtain very similar results with all major conclusions unchanged (keeping in mind that lnRR and Hedge's are on different scales):  
+ The effect of sexual selection on population fitness is generally positive.
+ This effect is magnified for females in stressful environments with a significantly positive interaction term (stress*females)

```{r, warning = F, fig.width=7, fig.height = 3.5, message = F}
grid.arrange(both.plots + 
               guides(fill = FALSE) + 
               ggtitle("Comparison between Hedge's g and lnRR effect sizes \n\n"), 
             lnRR.plots + ggtitle("\n\n"), nrow = 1)
```

For these reasons we have decided to use Hedge's g as the main effect size throughout the manuscript but reference and provide equivelant statistical analysis for lnRR in the supplementary material.

**2 --- The authors may consider also doing another set of meta-analyses using lnVR (log variability ratio - proposed in Nakagawa et al. 2015).**

**Nakagawa, S., R. Poulin, K. Mengersen, K. Reinhold, L. Engqvist, M. Lagisz, and A. M. Senior. 2015. Meta-analysis of variation: ecological and evolutionary applications and beyond. \textit{Methods in Ecology and Evolution} 6:143-152.**

Answer

**As one can see in Figure 2, the mean results (Hedges' $g$) are a mirror image of the variance results (this makes sense CV controls for means). I would like to see what the absolute change in variances. Probably the authors can put the analysis using lnVR in the supplement. The results of this analysis can be discussed. Also, the mean-variance relationship between mean and variance (sd) should be verified (e.g. plot log mean and log sd or log variance).**

The reviewer suggestions have been included in our meta-analysis in order to validate our conclusions. Several key points arise from the meta-analysis of variation:

*Firstly*, as the reviewer suggested we justify the use of lnCVR my assessing the mean-variance relationship. Given this finding the use of lnCVR is preferenced over lnVR, as lnVR does not account for the mean-variance relationship.

```{r, warning = F, message=F}
full_dataset %>% ggplot(aes(x = mean.high, y = sd.high))+
  geom_point()+
  scale_x_log10()+
  scale_y_log10()+
  xlab("Mean of Treatment Group")+
  ylab("Standard Deviation of Treatment Group") +
  geom_smooth(method = "lm", colour = "red")
```

*Secondly*, We conducted both Bayesian and REML meta-analyses for lnVR, which we present in the supplementary material and briefly discuss in the manuscript. We found similar trends for the meta-analysis of variance using lnCVR and the meta-analysis of variance using lnVR, These are:

+  In stressful environments sexual selection leads to a decrease in the variance for females, relative to benign environments
+  Bayesian models suggest that under stressful conditions elevated sexual selection causes a significant decrease in the variance. But this trend is not significant for REML models.

However, when using lnVR, the variability of fitness traits for males under sexual selection increased, while in the model using lnCVR it did not.

```{r, warning = F, fig.width=7, fig.height = 3.5, message = F}
grid.arrange(both.var.plots + 
               guides(fill = FALSE) + 
               ggtitle("Comparison between lnCVR and lnVR effect sizes \n\n"), 
             both.lnVR.plots + ggtitle("\n\n"), nrow = 1)
```

**3 --- $I^2$ needs to be explained. $I^2$ is proposed originally here:**

**Higgins, J., and S. Thompson. 2002. Quantifying heterogeneity in a meta-analysis. \textit{Statistics in Medicine} 21:1539 - 1558.**

**But later expanded in here for mixed models (hierarchical models)**

**Nakagawa, S., and E. S. A. Santos. 2012. Methodological issues and advances in biological meta-analysis. \textit{Evolutionary Ecology} 26:1253-1274.**

**This is what the authors use. Also, it will be good to put the degree in I^2 in the context. To do this see this paper:**

**Senior, A. M., C. E. Grueber, T. Kamiya, M. Lagisz, K. O'Dwyer, E. S. A. Santos, and S. Nakagawa. 2016. Heterogeneity in ecological and evolutionary meta-analyses: its magnitude and implications. \textit{Ecology} 97:3293-3299.**

Answer

**4 --- Publication tests have been conducted on, I think, "meta-analytic" residuals results as suggested in Nakagawa and Santos 2012. One could use such residuals to conduct the trim and fill method and see how much the mean could move (see Nakagawa and Santos 2012). One should remember the funnel asymmetry could be caused by the presece of heterogeniety. See:**

**Egger, M., G. Smith, M. Schneider, and C. Minder. 1997. Bias in metaanalysis detected by a simple, graphical test. \textit{Br Med J} 315:629 - 634.**

Answer

**5 --- Figure 3 - are these grey envelopes 95% CI?**

Answer

**6 --- the title - do the authors include "a systematic review" - so "a systematic reveiw and meta-analysis" - these two things are diferent - see:**

**Nakagawa, S., and R. Poulin. 2012. Meta-analytic insights into evolutionary ecology: an introduction and synthesis. \textit{Evolutionary Ecology} 26:1085-1099.**

Our mstudy does not involve a systematic review.

**Hope my comments are useful.**

They were extremely useful, many thanks!

________________________________________________________

## Comments from Reviewer 3

**This manuscript addresses the question of whether, on average, sexual selection has a net beneficial or detrimental effect on population fitness, using a meta-analysis of experimental evolution studies comparing the fitness of populations under different intensities of sexual selection. The analyses consider the effects of sexual selection on both the mean and variance of population fitness, and test whether these effects differ for stressful vs. benign environments, or depend on the measure of population fitness.**

**These are issues of longstanding interest, and as there is now quite a wealth of experimental evolution studies addressing these questions a meta-analysis synthesising their findings is timely.**

Many thanks for your time and attention, and for the valuable feedback.

**The results of this meta-analysis broadly concur with current theoretical predictions, and are likely to interest a wide readership. Across all studies, environment types and fitness measures there was a small positive effect of sexual selection on mean population fitness, although effect sizes varied with the fitness measure used. The positive effect of sexual selection on fitness was strongest for females in stressful environments, who also showed reduced variance in fitness under sexual selection.**

Answer

**The authors discuss these findings thoughtfully, although they focus more on the (weaker) effects found for ‘direct’ measures of fitness while somewhat neglecting the effects seen for ‘indirect’ fitness measures (e.g. of attractiveness, mating latency, lifespan, ejaculate traits). These potentially deserve more consideration, especially as many appear to be male-limited traits or measured mainly in males, that would appear to have a direct bearing on male mating success (if an indirect link to overall fitness), yet despite this the overall effect of sexual selection on male fitness is not significant.**

Need to answer this in a similar line to Jaceks comments.

**I have only a few more specific comments:**

**Does your approach consider variation in the intensity of sexual selection exclusively on males? This is worth noting (e.g. statements such as that on line 98-99 might be amended to “Sexual selection on males significantly improved female fitness.”)**

Answer

**Experiments manipulating the sex ratio to alter sexual selection might simultaneously decrease sexual selection on males while increasing sexual selection on females. Would this be classified as reduced sexual selection in your analyses? I don’t think you included this aspect of study design as a moderator variable in any of your analyses – given that close to half of the effects you include come from “alternative manipulations” (line 70) perhaps you could test whether this affects effect size?**

Answer

**It seems possible that sexual selection on females, and not only on males, could affect population fitness – potentially even more directly than sexual selection on males. And might differing extents of sexual selection in each sex interact with the differences you saw in fitness measured in males vs. females?**

Answer

**L111-118 It seems a little odd that there is no residual heterogeneity in your $I^2$ estimates. From the supplemental information it is not entirely clear to me how you have adapted the function that is under development in “metaAidR”, but can you double check that you have appropriately incorporated the residual variance into these estimates?**

Answer

**L307-311 Please clarify here that lnCVR was calculated as the ratio $ln(CVfitness_{SS}/CVfitness_{no SS})$**

Answer

**Fig. 3 Please explain the dashed red and black lines in the figure caption.**

Answer

**Tables 1, 2 Please check – the parameter estimates for ‘Female sex’ and test for ‘Female > Male’ are bolded inconsistently between the REML and Bayesian models where these estimates are identical.**

Answer
